### 2015-07-28 motivation.tex
Address shepherd comments through motivatoin.

\paragraph{Traditional package managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
Others, such as Anaconda~\cite{anaconda,conda}, use the same approach but
are designed to run on top of a host OS.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package only has a single version
and most of the tools install packages in a single, inflexible location.
To install multiple configurations, the user must create a custom, combinatorial
naming scheme to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

@@

To cope with software diversity, many HPC sites use a combination of existing
package managers and either manual or semi-automated installation.
For the baseline OS, many sites maintain traditional binary
packages using the vendor's package manager. LLNL maintains a Linux
distribution, CHAOS~\cite{chaos} for this purpose, which is managed using RPM.
The popular ROCKS~\cite{rocks} cluster distribution uses RPM and Anaconda
in a similar fashion.
%
For custom builds, many sites adhere to detailed naming conventions
that encode information in file system paths.
Table~\ref{tab:naming-conventions} shows several sites' conventions.
LLNL uses the APT package manager for installs
in the {\tt /usr/local/tools} file system and {\tt /usr/global/tools}
for manual installs.
ORNL uses hand installs but adheres to strict scripting conventions
to reproduce each build~\cite{jones+:cug08}.
TACC relies heavily on locally maintained RPMs.

@@

Diverse software versions not only present problems for build and installation;
they also complicate the runtime environment. When launched, an executable
must determine the location of its dependency libraries, or it will not run.
Even worse, it may find the wrong dependencies and subtly produce incorrect results.
Statically linked binaries do not have this issue, but modern
operating systems make extensive use of dynamic linking.
By default, the dynamic loader on most systems is configured to search only
system library paths such as {\tt /lib}, {\tt /usr/lib}, and
{\tt /usr/local/lib}.  If binaries are installed in other locations, the
{\it user} who runs the program must typically add dependency library paths to
{\tt LD\_LIBRARY\_PATH} (or a similar environment variable) so that the loader
can find them.  The user is often not the same person who installed the binary,
so determining what paths to add can be difficult, especially for inexperienced users.

Many HPC sites address this problem using {\it environment modules}, which
allow users to ``load'' and ``unload'' such settings dynamically using simple
commands. Environment modules emerged in 1991, and there are many implementations~\cite{dotkit,furlani+:lisa91,furlani+:lisa96,mclay:lmod,mclay:lmod-tutorial}.
The most advanced of these, such as Lmod~\cite{mclay:lmod,mclay:lmod-tutorial},
provide software hierarchies similar to the naming conventions in
Table~\ref{tab:naming-conventions}. They allow users to load a software stack
quickly if they know which one is required.

The alternative to per-user environment settings is to embed library search
paths in installed binaries at compile time. When set this way, the search
path is called an {\tt RPATH}. {\tt RPATHs} and environment modules are not
mutually exclusive. Modules can still be used to set {\tt MANPATH}, {\tt PATH},
etc., while adding {\tt RPATHs} ensures that binaries run correctly
regardless of whether the right module is loaded. LC installs software with
both {\tt RPATHs} and {\tt dotkit}~\cite{dotkit} modules.

@@

The EasyBuild~\cite{hoste+:pyhpc12} tool is in production use at
the University of Ghent and several other sites.  It allows multiple versions
to be installed at once.  Rather than setting {\tt RPATHs}, it
auto-generates module files
to manage each package's environment, and it is closely coupled with
Lmod~\cite{geimer+:hust14}.  EasyBuild groups the compiler, MPI, FFT, and
BLAS libraries together in a {\it toolchain} that can be combined with
any package file. This provides some degree of composability and
separates compiler flags and MPI concerns from client packages.

HashDist~\cite{hashdist} is a meta-build system and package manager
for HPC.  Of the existing solutions, it is the most similar to spack.
Like Nix, it uses cryptographic versioning and stores installations in
unique directories.
%
Both Nix and HashDist use {\tt RPATHs} in their packages to ensure that
libraries are found correctly.

@@

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatorial, and arbitrary combinations 
of compiler, MPI version, build options, and dependency versions requires 
tedious modifications of many package files. Indeed, the number of package 
files required for most existing systems scales with the number of version 
{\it combinations}, not the number of packages, which quickly becomes 
unmanageable.  As an example, the EasyBuild system has over 3,300 files for 
several permutations of only 600 packages. A slightly different dependency 
graph requires an entire new package file hierarchy.  HashDist supports 
composition more robustly but does not have first-class parameters for 
versions, compilers, or versioned interfaces.
%
HPC sites need better mechanisms to {\it parameterize} packages so that new
builds can be {\it composed} in response to user needs.
%

>>>>

\paragraph{Traditional package managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
Anaconda~\cite{anaconda,conda}, uses the same approach but
is designed to run on top of a host OS.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package has only a single version,
and most install packages in a single, inflexible location.
To install multiple configurations, users must create custom, combinatorial
naming schemes to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

@@

To cope with software diversity, many HPC sites use a combination of existing
package managers and either manual or semi-automated installation.
For the baseline OS, many sites maintain traditional binary
packages using the vendor's package manager. LLNL maintains a Linux
distribution, CHAOS~\cite{chaos} for this purpose, which is managed using RPM.
The popular ROCKS~\cite{rocks} cluster distribution uses RPM and Anaconda
in a similar fashion.
%
For custom builds, many sites adhere to detailed naming conventions
that encode information in file system paths.
Table~\ref{tab:naming-conventions} shows several sites' conventions.
LLNL uses the APT package manager for installs
in the {\tt /usr/local/tools} file system and {\tt /usr/global/tools}
for manual installs.
Oak Ridge National Laboratory (ORNL) uses hand installs but adheres
to strict scripting conventions
to reproduce each build~\cite{jones+:cug08}.
The Texas Advanced Computing Center (TACC) relies heavily on locally maintained RPMs.

@@

Diverse software versions not only present problems for build and installation;
they also complicate the runtime environment. When launched, an executable
must determine the location of its dependency libraries, or it will not run.
Even worse, it may find the wrong dependencies and subtly produce incorrect results.
Statically linked binaries do not have this issue, but modern
operating systems make extensive use of dynamic linking.
By default, the dynamic loader on most systems is configured to search only
system library paths such as {\tt /lib}, {\tt /usr/lib}, and
{\tt /usr/local/lib}.  If binaries are installed in other locations, the
{\it user} who runs the program must typically add dependency library paths to
{\tt LD\_LIBRARY\_PATH} (or a similar environment variable) so that the loader
can find them.  Often, the user is not the same person who installed the library,
and even advanced users may have difficulty determining which paths to add.

Many HPC sites address this problem using {\it environment modules}, which
allow users to ``load'' and ``unload'' such settings dynamically using simple
commands. Environment modules emerged in 1991, and there are many implementations~\cite{dotkit,furlani+:lisa91,furlani+:lisa96,mclay:lmod,mclay:lmod-tutorial}.
The most advanced of these, Lmod~\cite{mclay:lmod,mclay:lmod-tutorial},
provides software hierarchies similar to the naming conventions in
Table~\ref{tab:naming-conventions}. This allows users to load a software stack
quickly if they know which one is required.

The alternative to per-user environment settings is to embed library search
paths in installed binaries at compile time. When set this way, the search
path is called an {\tt RPATH}. {\tt RPATHs} and environment modules are not
mutually exclusive. Modules can still be used to set variables that are
unrelated to linking, such as {\tt MANPATH} and {\tt PATH}.  Adding
{\tt RPATHs} still ensures that binaries run correctly,
regardless of whether the right module is loaded. LC installs software with
both {\tt RPATHs} and {\tt dotkit}~\cite{dotkit} modules.

@@

The EasyBuild~\cite{hoste+:pyhpc12} tool is in production use at
the University of Ghent and the J\"ulich Supercomputing Center.
It allows multiple versions to be installed at once.  Rather
than setting {\tt RPATHs}, it auto-generates module files
to manage the environment, and it is closely coupled with
Lmod~\cite{geimer+:hust14}.  EasyBuild groups the compiler, MPI, FFT, and
BLAS libraries together in a {\it toolchain} that can be used by
package files. This provides some degree of composability and
separates compiler flags and MPI concerns from client packages.

HashDist~\cite{hashdist} is a meta-build system and package manager
for HPC.  Of the existing solutions, it is the most similar to Spack.
Like Nix, it uses cryptographic versioning and stores installations in
unique directories.
%
Both Nix and HashDist use {\tt RPATHs} in their packages to ensure that
libraries are found correctly.

@@

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatorial, and arbitrary combinations 
of compiler, MPI version, build options, and dependency versions require 
non-trivial modifications to many package files. Indeed, the number of package 
files required for most existing systems scales with the number of version 
{\it combinations}, not the number of packages, which quickly becomes 
unmanageable.  As an example, the EasyBuild system has over 3,300 files for 
several permutations of only 600 packages. A slightly different dependency 
graph requires an entire new package file hierarchy.  HashDist supports 
composition more robustly but does not have first-class parameters for 
versions, compilers, or versioned interfaces.
%
HPC sites need better ways to {\it parameterize} packages so that new
builds can be {\it composed} in response to user needs.
%
